\vspace{1.0cm} \LARGE \bf Structured population models demystify MERS-CoV epidemiology

### 2017-07-12 mers-structure.tex
Add algorithm float.

\usepackage{longtable}

@@

As of 10 May 2017, the World Health Organization has been notified of 1952 cases of MERS-CoV.
We thus simulate final transmission chain sizes using equation \ref{clusterLikelihood} \citep{lloyd-smith_superspreading_2005,blumberg_inference_2013} until we reach an epidemic comprised of 2000 cases.
10 000 simulations were run for 121 uniformly spaced values of R$_{0}$ across the range [0.5--1.1] with dispersion parameter $\omega$ fixed to 0.1.
Each simulation results in a vector of outbreak sizes $C$, where $C_{i}$ is the size of the \textit{i}th transmission cluster and $\sum_{i=1}^{K} C_{i} = 2000$, where $K$ is the number of clusters generated.
We sample from the case cluster size vector according to a multivariate hypergeometric distribution to simulate sequencing.
The resulting sequence cluster size vector $S$ contains $K$ entries, some of which are zero (\textit{i.e.} case clusters not sequenced), but $\sum_{i=1}^{K} S_{i} = 174$ which reflects the number of MERS-CoV sequences used in this study.
Since the sampling scheme operates under equi-probable sequencing we also simulated biased sequencing by using concentrated hypergeometric distributions where the probability mass function is squared (bias = 2) or cubed (bias = 3) and then normalized.
This makes clusters likely to be `sequenced' even more likely to be sequenced and \textit{vice versa}.
We performed a smaller set of simulations with 2500 replicates and twice the number of cases, \textit{i.e.} $\sum_{i=1}^{K} C_{i} = 4000$, to explore a dramatically underreported epidemic.

@@

>>>>

\usepackage{longtable}
\usepackage[]{algorithm2e}

@@

As of 10 May 2017, the World Health Organization has been notified of 1952 cases of MERS-CoV.
We thus simulate final transmission chain sizes using equation \ref{clusterLikelihood} \citep{lloyd-smith_superspreading_2005,blumberg_inference_2013} until we reach an epidemic comprised of 2000 cases.
10 000 simulations were run for 121 uniformly spaced values of R$_{0}$ across the range [0.5--1.1] with dispersion parameter $\omega$ fixed to 0.1.
Each simulation results in a vector of outbreak sizes $C$, where $C_{i}$ is the size of the \textit{i}th transmission cluster and $\sum_{i=1}^{K} C_{i} = 2000$, where $K$ is the number of clusters generated.
We sample from the case cluster size vector according to a multivariate hypergeometric distribution to simulate sequencing (algorithm \ref{hypergeometric}).
The resulting sequence cluster size vector $S$ contains $K$ entries, some of which are zero (\textit{i.e.} case clusters not sequenced), but $\sum_{i=1}^{K} S_{i} = 174$ which reflects the number of MERS-CoV sequences used in this study.
Since the sampling scheme operates under equi-probable sequencing we also simulated biased sequencing by using concentrated hypergeometric distributions where the probability mass function is squared (bias = 2) or cubed (bias = 3) and then normalized.
This makes clusters likely to be `sequenced' even more likely to be sequenced and \textit{vice versa}.
We performed a smaller set of simulations with 2500 replicates and twice the number of cases, \textit{i.e.} $\sum_{i=1}^{K} C_{i} = 4000$, to explore a dramatically underreported epidemic.

@@

\begin{algorithm}[H]
 \KwData{Array of case cluster sizes in outbreak $[C_{i},C_{i+1},C_{i+2}...C_{K}]$, sequences available $S$, total outbreak size $\sum_{i=1}^{K} C_{i}$}
 \KwResult{Array of sequence cluster sizes sampled: $[S_{i},S_{i+1},S_{i+2}...S_{K}]$}
 Draw from $C_{i}$ according to a hypergeometric distribution with $C_{i}$ successes, $\sum_{i=1}^{K} C_{i}-C_{i}$ failures after $S$ trials\;
 \While{$i < K$}{
  $i = i + 1$\;
  $S = S - S_{i-1}$\;
  Recover probability mass function $p = [p(0), ..., p(C_{i})]$ for a hypergeometric distribution with $C_{i}$ successes, $\sum_{i=1}^{K} C_{i}-C_{i}$ failures after $S$ trials\;
  (Optional) Introduce bias into $p_{bias} = \frac{[p(0)^{bias}, ..., p(C_{i})^{bias}]}{\sum p_{bias}}$\;
  Draw a sequence cluster size $S_{i}$ from array of potential sequence cluster sizes $[0, ..., C_{i}]$ according to $p$\;
 }
 Add remaining sequences to last sequence cluster $C_{K} = S - S_{K-1}$\;
 \caption{\textbf{Multivariate hypergeometric sampling scheme.}
 Pseudocode describes the multivariate hypergeometric sampling scheme that simulates sequencing.
 Probability of sequencing a given number of cases from a case cluster depends on cluster size and sequences left (\textit{i.e.} ``sequencing capacity'').
 The bias parameter determines how probability mass function of the hypergeometric distribution is concentrated.
 }
 \label{hypergeometric}
\end{algorithm}
